---
title: "PREDICT_2019"
author: "Ethan Greig"
date: "11/17/2018"
output: pdf_document
---

## Import Raw Team Data

# import from morehockeystats: ENGF, ENGA, B-S stats, Time_Led (set custom format h:mm, calc *1440, /60)
#   *move WPG for ATL years, and move ARI for 2015
# import from natural stat trick: W, ROW, HDCF%_55
# import from Fox Sports: PP%, PK% (paste as text, text->columns w/ tab delim)
```{r}
corsica_55 <- read.csv("input_csvs/corsica_5v5_08-18.csv", check.names=FALSE)
corsica_55$Season <- strtoi(substr(corsica_55$Season, 6, 9))
corsica_55 <- corsica_55[, c('Team', 'Season', 'CF%_55', 'CF/60_55', 'CA/60_55', 'GF%_55', 'GF/60_55', 'GA/60_55', 'xGF%_55', 'xGF/60_55', 'xGA/60_55', 'Sh%_55', 'Sv%_55')]

corsica_all <- read.csv("input_csvs/corsica_all_08-18.csv", check.names=FALSE)
corsica_all$Season <- strtoi(substr(corsica_all$Season, 6, 9))
corsica_all <- corsica_all[, c('Team', 'Season', 'CF%_all', 'CF/60_all', 'CA/60_all', 'GF_all', 'GA_all', 'GF%_all', 'GF/60_all', 'GA/60_all', 'xGF%_all', 'xGF/60_all', 'xGA/60_all', 'Sh%_all', 'Sv%_all')]

other_data <- read.csv("input_csvs/misc_stats_08-18.csv", check.names=FALSE)
other_data <- other_data[, c('Team', 'Season', 'HDCF%_55', 'Time_Led', 'ENGF_all', 'ENGA_all', 'PP%', 'PK%', 'W', 'ROW', 'Berg', 'xBerg')]

stats <- c('CF%_55', 'CF/60_55', 'CA/60_55', 'GF%_55', 'GF/60_55', 'GA/60_55', 'xGF%_55', 'xGF/60_55', 'xGA/60_55', 'Sh%_55', 'Sv%_55', 'CF%_all', 'CF/60_all', 'CA/60_all', 'GF%_all', 'ENGD', 'GF/60_all', 'GA/60_all', 'xGF%_all', 'xGF/60_all', 'xGA/60_all', 'Sh%_all', 'Sv%_all', 'HDCF%_55', 'Time_Led', 'PP%', 'PK%', 'W', 'ROW', 'Berg', 'xBerg')

raw_stats <- merge(merge(corsica_55, corsica_all, by=c("Team", "Season")), other_data, by=c("Team", "Season"))
raw_stats$`ENGD` <- with(raw_stats, ENGF_all-ENGA_all)
raw_stats <- raw_stats[, c('Team', 'Season', stats)]
raw_stats
```


## Create Matchups
```{r}
dStats <- stats
for (i in 1:length(stats)) {
  stat <- stats[i]
  dStat <- paste0('d', stat)
  dStats[i] <- dStat
}

dStats <- make.names(dStats)

# Import series data, add empty columns for stat differences
series <- read.csv("input_csvs/all_playoff_series.csv")
series = series[series$Season<2018,]
series$Home_Won <- round(series$Home_W., 0)
for(i in 1:length(dStats)) {
  dStat <- dStats[i]
  series[,dStat] <- NA
}

# define function to extract stat differences given two teams and a season
getDifferences <- function(row, df) {
  df_year <- df[df$Season==row$Season,]
  home_team <- df_year[as.character(df_year$Team)==as.character(row$Home),]
  away_team <- df_year[as.character(df_year$Team)==as.character(row$Away),]
  for (i in 1:length(stats)) {
    stat <- stats[i]
    dStat <- dStats[i]
    row[dStat] <- home_team[stat] - away_team[stat]
  }
  return(row)
}

# fill series data frame with stat differences from regular season
for (row in 1:nrow(series)) {
  series[row,] <- getDifferences(series[row,], raw_stats)
}

series
```

# Remove highly-correlated features
```{r}
corr <- cor(series[,c('Home_Won', dStats)])
corr_vars <- names(data.frame(corr))[which(corr>0.85 & corr<1, arr.ind = TRUE)[, "col"]]
cor(series[,c('Home_Won', corr_vars)])

# Judgement call on which of corr pairs to remove
dStats <- dStats[!dStats%in%c('dCF._55', 'dCF.60_all', 'dCA._all', 'dBerg')]
```

## Define Some Metrics
```{r}
logLoss <- function(pred, actual){
  -1*mean(log(pred[model.matrix(~ actual + 0) - pred > 0]))
}

acc <- function(pred, actual){
  wlpred <- round(pred, 0)
  correct <- wlpred==actual
  num_correct <- sum(correct, na.rm=TRUE)
  return(num_correct/length(correct))
}
```

## Symmetric Data Frame
```{r}
df <- series[,c('Home_Won', dStats)] # keep only relevant numerical rows. Home_Adv?
df_flip <- -df
df_flip$Home_Won <- 1-df$Home_Won
df_symm <- rbind(df, df_flip)
df_symm[,dStats] <- data.frame(scale(df_symm[,dStats]))
```

## Cross-Validation
```{r}
library(randomForest)

cv <- function(data, model='log', nfolds=5, nruns=150) {
  n <- nrow(data)
  results <- matrix(nrow=nruns, ncol=2)
  
  for (i in 1:nruns) {
    data_shuffled <- data[sample(n),]
    folds <- cut(seq(1,n),breaks=nfolds,labels=FALSE)
    
    log_loss <- NA
    accuracy <- NA
    
    for (j in 1:nfolds) {
    test_rows <- which(folds==j,arr.ind=TRUE)
    test_data <- data_shuffled[test_rows,]
    train_data <- data_shuffled[-test_rows,]
    
    if (model=='log') {
      fold_model <- suppressWarnings(glm(`Home_Won` ~ .,data=train_data, family=binomial(link='logit')))
    }
    else if (model=='rf') {
      fold_model <- suppressWarnings(randomForest(`Home_Won` ~ .,data=train_data))
    }
    
    test_data$fold_prediction <- predict(fold_model, test_data[,dStats], type='response')
  
    log_loss[j] <- logLoss(test_data$fold_prediction, test_data$Home_Won)
    accuracy[j] <- acc(test_data$fold_prediction, test_data$Home_Won)
    }
    
  results[i,] = c(mean(log_loss), mean(accuracy))
  }
  
return(colMeans(results))
}
```

## Benchmark Models
```{r}
cv(df, nfolds=5)
cv(df_symm, nfolds=5)
cv(df, nfolds=10)
cv(df_symm, nfolds=10)
```
