---
title: "predict_2021"
author: "Ethan Greig"
date: "5/16/2021"
output: pdf_document
---
## Import Raw Team Data

# With Corsica shutting down, data must be imported from a slightly different CSV format for 2019 onward

# import from evolving hockey: corsica stats + copy/paste Sh% and Sv% from past years corsica or current naturalstattrick
# import from morehockeystats: ENGF, ENGA, B-S stats, Time_Led (set to number format, multiply by 24)
# import from nhl.com: PP% and PK% (export to csv)
#   *move WPG for ATL years, and move ARI for 2015
# import from naturalstattrick: W, ROW, HDCF%_55
```{r}
library(knitr)
source(purl('import_and_generate_series.Rmd', output = tempfile()))
raw_stats
series
```

```{r}
dStats_4 <- c("dCF.60_55", "dGF._55", "dxGF._all",  "dPK.")
```

```{r}
library(randomForest)

predict_round <- function(series_data, year, round=0) {
  prior_series <- series_data[series_data$Season<year | (series_data$Season==year & series_data$Round<round),]
  train <- prior_series[,c('Home_Won', dStats)]

  if (round==0) pred_series <- series_data[series_data$Season==year,]
  else pred_series <- series_data[series_data$Season==year & series_data$Round==round,]
  
  rf_4 <- suppressWarnings(randomForest(`Home_Won` ~ ., data=train[,c('Home_Won', dStats_4)], ntree=500000, mtry=1, nodesize=25))
  
  output <- pred_series[,c('Season', 'Round', 'Home', 'Away')]
  output$rf_4_pred <- predict(rf_4, pred_series, type='response')

  write.table(output, file="~/repos/nhl-playoff-model/output_csvs/predictions_2021_temp.csv", sep=',', col.names=NA)
}
```

```{r}
predict_round(series, 2021, 0)
```