---
title: "predict_2019"
author: "Ethan Greig"
date: "4/7/2019"
output: pdf_document
---
## Import Raw Team Data

# With Corsica shutting down, data must be imported from a slightly different CSV format for 2019

# import from evolving hockey: corsica stats + copy/paste Sh% and Sv% from past years corsica or current naturalstattrick
# import from morehockeystats: ENGF, ENGA, B-S stats, Time_Led (set custom format h:mm, calc *1440, /60)
#   *move WPG for ATL years, and move ARI for 2015
# import from naturalstattrick: W, ROW, HDCF%_55
# import from Fox Sports: PP%, PK% (paste as text, text->columns w/ tab delim)
```{r}
EH_55 <- read.csv("input_csvs/EH_5v5_08-19.csv", check.names=FALSE)
EH_55$Season <- strtoi(substr(EH_55$Season, 5, 8))
EH_55 <- EH_55[, c('Team', 'Season', 'CF.60_55', 'CA.60_55', 'GF.60_55', 'GA.60_55', 'xGF.60_55', 'xGA.60_55', 'Sh%_55', 'Sv%_55')]
EH_55$`CF%_55` <- round(100* EH_55$`CF.60_55`/(EH_55$`CF.60_55` + EH_55$`CA.60_55`), 2)
EH_55$`GF%_55` <- round(100* EH_55$`GF.60_55`/(EH_55$`GF.60_55` + EH_55$`GA.60_55`), 2)
EH_55$`xGF%_55` <- round(100* EH_55$`xGF.60_55`/(EH_55$`xGF.60_55` + EH_55$`xGA.60_55`), 2)

EH_all <- read.csv("input_csvs/EH_all_08-19.csv", check.names=FALSE)
EH_all$Season <- strtoi(substr(EH_all$Season, 5, 8))
EH_all <- EH_all[, c('Team', 'Season', 'CF.60_all', 'CA.60_all', 'GF.60_all', 'GA.60_all', 'xGF.60_all', 'xGA.60_all', 'Sh%_all', 'Sv%_all')]
EH_all$`CF%_all` <- round(100* EH_all$`CF.60_all`/(EH_all$`CF.60_all` + EH_all$`CA.60_all`), 2)
EH_all$`GF%_all` <- round(100* EH_all$`GF.60_all`/(EH_all$`GF.60_all` + EH_all$`GA.60_all`), 2)
EH_all$`xGF%_all` <- round(100* EH_all$`xGF.60_all`/(EH_all$`xGF.60_all` + EH_all$`xGA.60_all`), 2)

other_data <- read.csv("input_csvs/misc_stats_08-19.csv", check.names=FALSE)
other_data <- other_data[, c('Team', 'Season', 'HDCF%_55', 'Time_Led', 'ENGF', 'ENGA', 'PP%', 'PK%', 'W', 'ROW', 'Berg')]
other_data$ENGD <- other_data$ENGF - other_data$ENGA

stats <- c('CF%_55', 'CF.60_55', 'CA.60_55', 'GF%_55', 'GF.60_55', 'GA.60_55', 'xGF%_55', 'xGF.60_55', 'xGA.60_55', 'Sh%_55', 'Sv%_55', 'CF%_all', 'CF.60_all', 'CA.60_all', 'GF%_all', 'ENGD', 'GF.60_all', 'GA.60_all', 'xGF%_all', 'xGF.60_all', 'xGA.60_all', 'Sh%_all', 'Sv%_all', 'HDCF%_55', 'Time_Led', 'PP%', 'PK%', 'W', 'ROW', 'Berg')

raw_stats <- merge(merge(EH_55, EH_all, by=c("Team", "Season")), other_data, by=c("Team", "Season"))
raw_stats <- raw_stats[, c('Team', 'Season', stats)]
raw_stats
```

## Create Matchups
```{r}
dStats <- stats
for (i in 1:length(stats)) {
  stat <- stats[i]
  dStat <- paste0('d', stat)
  dStats[i] <- dStat
}

dStats <- make.names(dStats)

# Import series data, add empty columns for stat differences
series <- read.csv("input_csvs/all_playoff_series.csv")
series$Home_Won <- round(series$Home_W., 0)
for(i in 1:length(dStats)) {
  dStat <- dStats[i]
  series[,dStat] <- NA
}

# define function to extract stat differences given two teams and a season
getDifferences <- function(row, df) {
  df_year <- df[df$Season==row$Season,]
  home_team <- df_year[as.character(df_year$Team)==as.character(row$Home),]
  away_team <- df_year[as.character(df_year$Team)==as.character(row$Away),]
  for (i in 1:length(stats)) {
    stat <- stats[i]
    dStat <- dStats[i]
    row[dStat] <- home_team[stat] - away_team[stat]
  }
  return(row)
}

# fill series data frame with stat differences from regular season
for (row in 1:nrow(series)) {
  series[row,] <- getDifferences(series[row,], raw_stats)
}

series
```

```{r}
dStats_5 <- c("dCF.60_55", "dGF._55","dGF._all", "dxGF._all",  "dPK.")
```

```{r}
library(randomForest)

predict_round <- function(series_data, year, round=0) {
  prior_series <- series_data[series_data$Season<year | (series_data$Season==year & series_data$Round<round),]
  train <- prior_series[,c('Home_Won', dStats)]

  if (round==0) pred_series <- series_data[series_data$Season==year,]
  else pred_series <- series_data[series_data$Season==year & series_data$Round==round,]
  
  rf_5 <- suppressWarnings(randomForest(`Home_Won` ~ ., data=train[,c('Home_Won', dStats_5)], ntree=50000, mtry=1, nodesize=107))
  
  output <- pred_series[,c('Season', 'Round', 'Home', 'Away')]
  output$rf_5_pred <- predict(rf_5, pred_series, type='response')

  write.table(output, file="~/repos/nhl-playoff-model/output_csvs/predictions_2019_temp.csv", sep=',', col.names=NA)
}
```

```{r}
predict_round(series, 2019, 0)
```